---
title: "Exploratory Data Analysis"
subtitle: "PSTAT 131"
author: "Isabella Escamilla"
date: "3/18/23"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
---

<style type = "text/css">
  body{
  font-family: trebuchet ms, sans-serif;
}
.list-group-item.active, .list-group-item.active:focus, .list-group-item.active:hover {
    background-color: #8A5667;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, eval = TRUE)
library(tidyverse)
library(ggplot2)
library(extrafont)
library(showtext)
library(gt)
library(grid)
library(gridExtra)

```

```{r, include=FALSE}
loadfonts(device='win', quiet = FALSE)
font_add('trebuchet', 'C:/windows/fonts/trebuc.ttf')
showtext_auto()
```

```{r, include=FALSE}
theme_polyvore <- function(){
  font <- 'trebuchet'
  
  theme_minimal() %+replace%
    
    theme(
      # grid elements
      #panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      axis.ticks = element_blank(),
      axis.line = element_line(),
      
      # text elements
      plot.title = element_text(
        family = font,
        size = 28,
        face = 'bold',
        hjust = 0,
        vjust = 2,
        margin = margin(b=5, t=5)),
      
      plot.subtitle = element_text(          #subtitle
                   family = font,            #font family
                   size = 14),               #font size
      
      plot.caption = element_text(           #caption
                   family = font,            #font family
                   size = 9,                 #font size
                   hjust = 1),               #right align
      
      axis.title = element_text(             #axis titles
                   family = font,            #font family
                   size = 20),               #font size
      
      axis.text = element_text(              #axis text
                   family = font,            #axis family
                   size = 16),                #font size
      
      axis.text.x = element_text(            #margin for axis text
                    margin=margin(5, b = 10)),
      
      axis.title.y = element_text(margin=margin(b=10, r=10), angle=90)
      

    )
}
```


## Exploratory Data Analysis
Now that we've tidied the data in a format that will be better suited for our analysis, we can begin by making some plots and seeing how the data is distributed.

First we will import the data and convert the appropriate variables into factors.

```{r}
# Reading the tidied data into R
outfits <- read.csv('tidy_data.csv')

# converting all character columns into factors
outfits[sapply(outfits, is.character)] <- lapply(outfits[sapply(outfits, is.character)], as.factor)
# I will only be using set_id to retrieve photos, so we will leave it be

outfits %>% head(5)
```


In the final data set, there are 14 variables:

Name | Variable description
---|---
`views` | The number of times an outfit has been viewed
`likes` | The number of likes an outfit received **(the response variable)**
`set_id` | identifies the particular outfit
`item_1` - `item_8` | gives the name of the item type for up to 8 items in an outfit
`color_1` - `color_4` | gives the unique colors found in the outfit for up to 4 different colors


### Likes
We will start by looking at the distribution of our response variable, `likes`.


```{r}
outfits %>% 
  ggplot(aes(x = likes)) + 
  geom_histogram(binwidth = 100, fill="#2C514C") +
  labs(x = 'Number of Likes', y = 'Count', title = 'Distribution of the Number of Likes') +
  coord_cartesian(xlim = c(0, 4000)) +
  theme_polyvore()
```

The number of likes ranges from 0 to around 3000, and the frequency decreases exponentially as the number of likes goes up. This follows the sort of logic you'd expect, where a select few get lots of likes while most go unseen.

### Items (Clothing Articles)
After an initial survey of the raw data, I noted that there were a lot of different item ids.

```{r, class.source = 'fold-show'}
# Finds the number of unique items in item_1 column
outfits$item_1 %>% unique() %>% length()
```
And indeed, there are 229 unique items in `item_1` alone! To further examine the nature of the item columns, we will look at a plot of the item frequencies in `item_1`. Additionally, we'll do a "zoom in" of the top 10 items by frequency in `item_1`.

```{r}
outfits %>% ggplot(aes(x = fct_infreq(item_1))) +
  geom_bar(fill = '#F08080') +
  labs(x = 'Item 1', y='', title = 'Frequencies of item_1 Items') +
  theme_polyvore() +
  theme(axis.text.x = element_blank(), 
        panel.grid.major.x = element_blank(), 
        axis.line = element_blank()) -> p1

item1_t10 <- (sort(table(outfits$item_1), decreasing = TRUE) %>% names)[1:10]
outfits_item1_top10 <- outfits[outfits$item_1 %in% item1_t10,]

outfits_item1_top10 %>% 
  ggplot(aes(y = fct_infreq(item_1))) +
  geom_bar(fill="#583742") +
  labs(x = '', y='', title = 'Top 10 item_1 Items') +
  theme_polyvore() +
  theme(axis.line = element_blank()) -> p2

grid.arrange(p1, p2, ncol=2)
```

The first plot shows that occurrences of the 229 items in `item_1` are concentrated towards the top few items and fall off significantly after. We can see from the second plot that the top items of `item_1` are all tops, dresses, and jackets of some type. This similarity gives us an idea of how the items have been organized, and we can get a better picture from constructing a table.

```{r}
item1_t5 <- (sort(table(outfits$item_1), decreasing = TRUE) %>% names)[1:5]
item2_t5 <- (sort(table(outfits$item_2), decreasing = TRUE) %>% names)[1:5]
item3_t5 <- (sort(table(outfits$item_3), decreasing = TRUE) %>% names)[1:5]
item4_t5 <- (sort(table(outfits$item_4), decreasing = TRUE) %>% names)[1:5]
item5_t5 <- (sort(table(outfits$item_5), decreasing = TRUE) %>% names)[1:5]
item6_t5 <- (sort(table(outfits$item_6), decreasing = TRUE) %>% names)[1:5]
item7_t5 <- (sort(table(outfits$item_7), decreasing = TRUE) %>% names)[1:5]
item8_t5 <- (sort(table(outfits$item_8), decreasing = TRUE) %>% names)[1:5]

item_freq_table <- data.frame(item_1 = item1_t5,
                              item_2 = item2_t5,
                              item_3 = item3_t5,
                              item_4 = item4_t5,
                              item_5 = item5_t5,
                              item_6 = item6_t5,
                              item_7 = item7_t5,
                              item_8 = item8_t5)

```

```{r}
item_freq_table %>% gt() %>% 
  tab_header(
    title = "Top 5 Items by Frequency"
  )
```

From looking at both the raw data and this table, a general pattern emerges: within each outfit, the items are sorted from most to least significant, with `item_1` being the most significant item and `item_8` being the least. The top items in `item_2` are still significant, with jackets being the most frequent and "bottoms" being common elements as well (shorts, jeans, skirts). `item_3` dips into shoe territory, and `item_4` through `item_8` can be seen mostly as the "accessory section". Another thing that is interesting to note from the table is that from `item_5` and onward, "None" is the most frequent item. We can check the proportion of "None"s pretty easily.

```{r}
# Gets proportion of data that has fewer than 5 items, fewer than 6 items, fewer than 7 items, and fewer than 8 items
props <- c(((filter(outfits, item_4 == 'None') %>% count()) / (outfits %>% count()))[1,1] %>% as.numeric(),
((filter(outfits, item_5 == 'None') %>% count()) / (outfits %>% count()))[1,1] %>% as.numeric(),
((filter(outfits, item_6 == 'None') %>% count()) / (outfits %>% count()))[1,1] %>% as.numeric(),
((filter(outfits, item_7 == 'None') %>% count()) / (outfits %>% count()))[1,1] %>% as.numeric(),
((filter(outfits, item_8 == 'None') %>% count()) / (outfits %>% count()))[1,1] %>% as.numeric())
props
```

With some simple arithmetic, this gives us a breakdown of the item number proportions:

* 8% of the outfits have 4 items,
* 15% have 5 items,
* 17% have 6 items,
* 19% have 7 items, and
* 41% have 8 items

Now we want to see what relationship the items have with likes. Before doing so, I will transform the data so that we only have 10 different categories, which will be the top 9 occurrences plus another category, 'other'.

```{r}
outfits2 <- outfits %>% mutate(item_1 = fct_lump_n(outfits$item_1, n=9))
colors <- c('#335C57', '#583742', '#F08080',
             '#335C57', '#583742', '#F08080',
             '#335C57', '#583742', '#F08080',
             '#335C57')
outfits2 %>% 
  ggplot(aes(x=likes, y=item_1, group=item_1, alpha=0.9)) +
  geom_boxplot(fill = colors) +
  scale_x_continuous(limits = c(0, 600)) +
  theme_polyvore() +
  theme(legend.position = 'none')
```

I set the y limit for `likes` at 600 so that the boxes aren't smushed around 0 (due to the nature of how `likes` is distributed). We can see that the items that yield more likes are sweaters and day dresses, while tank pops are the least popular of the category. 

### Colors
All of us understand that color plays an essential role in whether or not an outfit "works" (unless you're doing wardrobe for a 1947 film noir), so it will be interesting to see what kind of colors are represented in the data.

```{r}
colors1 <- c('#fb5c5c', '#4a687d', '#ffdfbf',
            'gray16', '#9b938c', '#e2ba59', '#628d55',
            '#643b34', '#e3774c', '#6e4258')

outfits %>% 
  ggplot(aes(y = fct_infreq(color_1))) +
  geom_bar(fill = colors1) +
  labs(x = '', y = '', title = 'Frequencies of Colors in color_1') +
  theme_polyvore() +
  theme(axis.line = element_blank(),
        panel.grid.major = element_blank()) -> p1

colors2 <- c('gray16', '#4e4c5e', '#ffdfbf', '#e2ba59',
             '#9b938c', '#643b34', '#4a687d', '#628d55',
             '#fb5c5c', '#e3774c', '#6e4258')

outfits %>% 
  ggplot(aes(y = fct_infreq(color_2))) +
  geom_bar(fill = colors2) +
  labs(x = '', y = '', title = 'Frequencies of Colors in color_2') +
  theme_polyvore() +
  theme(axis.line = element_blank(),
        panel.grid.major = element_blank()) -> p2

colors3 <- c('#4e4c5e', 'gray16', '#ffdfbf', '#e2ba59',
             '#9b938c', '#643b34', '#4a687d', '#6e4258',
             '#fb5c5c', '#628d55', '#e3774c')

outfits %>% 
  ggplot(aes(y = fct_infreq(color_3))) +
  geom_bar(fill = colors3) +
  labs(x = '', y = '', title = 'Frequencies of Colors in color_3') +
  theme_polyvore() +
  theme(axis.line = element_blank(),
        panel.grid.major = element_blank()) -> p3

colors4 <- c('#4e4c5e', 'gray16', '#9b938c', '#fb5c5c',
             '#e2ba59', '#4a687d', '#ffdfbf', '#6e4258',
             '#e3774c', '#643b34', '#628d55')

outfits %>% 
  ggplot(aes(y = fct_infreq(color_4))) +
  geom_bar(fill = colors4) +
  labs(x = '', y = '', title = 'Frequencies of Colors in color_4') +
  theme_polyvore() +
  theme(axis.line = element_blank(),
        panel.grid.major = element_blank()) -> p4

grid.arrange(p1, p2, p3, p4, ncol=2)

```

The color columns are similar to the item columns in the way they are ordered: `color_1` will have higher importance than `color_2` (and so on) in the outfit due to the way the information was extracted. This gives us some insight into the colors represented in these charts.

Red clearly reigns supreme for `color_1`, followed by blue, white, and black. We start to see "none" dominate starting in `color_2`, indicating that there are lots of outfits for which only one color could be extracted (or simply had a monochromatic scheme going on). In `color_4`, we see that most of the outfits simply don't have a color entry. This could mean that most outfits stuck to 1-3 colors.

Next, we'll take a look at the relationship between color and likes. 

```{r}
colors <- c('gray16', '#4a687d', '#643b34',
            '#9b938c', '#628d55', '#e3774c', '#6e4258',
            '#fb5c5c', '#ffdfbf', '#e2ba59')
            
outfits %>% 
  ggplot(aes(x=likes, y=color_1, group=color_1)) +
  geom_boxplot(fill=colors) +
  theme_polyvore()
```

All of the colors have their distribution of likes concentrated around 0 which makes sense when considering the distribution of the number of likes. The main differences we see come from the spread 

### Relationship between Items and Likes
Now that we have a good idea of what the different item columns hold, we can examine the items' relationship with the number of likes an outfit receives.




