---
title: "Data Exploration"
author: "Isabella Escamilla"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE, eval = FALSE)
library(tidyverse)
library(jsonlite)
library(httr)
library(rjson)
library(jsonlite)
```

### Importing the Data Set
To start off, we need to take a look at the data set and explore what it has to offer. Upon first look, we can see that the data comes in a nested format, with each outfit having the attributes `name`, `views`, `items`, `image`, `likes`, `date`, `set_url`, and `desc`. The `items` attribute is nested, which means it has its own set of attributes within it. This includes `index`, `name`, `price`, `likes`, `image`, and `categoryid`, which all describe parts of an outfit. The nesting isn't too bad, and the jsonlite package has some helpful functions which I will be using for the set.

```{r, eval=FALSE}
trainloc <- '/Users/bellaesc/Desktop/school/Year 3/Winter/PSTAT131/Project/polyvore_mlproject/polyvore_data/train_no_dup.json'
#trainloc <- './polyvore_data/train_no_dup.json'
train_json <- jsonlite::fromJSON(trainloc, flatten=TRUE, simplifyDataFrame = TRUE)
train_df <- as.data.frame(train_json)
```

```{r}

```


```{r}
dim(train_df)
head(train_df,n=3)
```

### Formatting the Data Frame
In the resulting data frame, each observation constitutes one outfit as intended, but the values in the `items` column are data frames. We can access these items and append them as new columns in the main data frame. We can start by examining the nests...

```{r}
train_df$items %>% head(1)
```

From each outfit's `items` data frame, we will want to take `name`, `price`, `likes`, and `categoryid`. The most simplest columns to add to the final data frame are `price` and `likes`, so we will start there.

```{r}
itemslist <- train_df$items
itemslist %>% head()
```
We will need to be able to perform an operation on a tibble of dataframes

*Note to future me:* `name` and `categoryid` carry similar information, so it will be easy to combine these two into one variable using PCA so that there are only up to 8 outfit variables instead of 16

https://stackoverflow.com/questions/43080531/r-expand-column-of-dataframes-in-existing-dataframe

### Trying to unpack the dataframe column (items)

```{r}
#as_tibble(train_df$items) %>% unnest() #error: missing values
```

* I'm clearly struggling. I'm gonna try this in python lol.


